---
title: "Documenting data and datasets"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Documenting data and datasets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

<!--
- TODO: Add vignette using documentation functions
-->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(qtkit)
```

: Data origin information

| Attribute                | Description                                               |
|--------------------------|-----------------------------------------------------------|
| Resource name            | The name of the resource.                                 |
| Data source              | URL, DOI, etc.                                            |
| Data sampling frame      | Language, language variety, modality, genre, etc.         |
| Data collection date(s)  | The dates the data was collected.                         |
| Data format              | .txt, .csv, .xml, .html, etc.                             |
| Data schema              | Relationships between data elements: files, folders, etc. |
| License                  | CC BY, CC BY-SA, etc.                                     |
| Attribution              | Citation information.                                     |

The {qtkit} package provides the function `create_data_origin()` to help create data origin files. By default the function will create the scaffolding for the data origin file in CSV format.


```{r}
ls("package:qtkit")
```
